Black swans in ecological time series\\ Black swans in population dynamics\\ Heavy-tailed ecological time series\\ Black swans in ecological dynamics\\ Black swan ecological dynamics\\ \ldots

### 2014-10-19 ms.tex
Add some notes for last paragraph

Black swans are unexpected extreme events with dramatic consequences. One of
the most striking black swans in ecology was the asteroid marking the mass
extinctions of the K-T boundary (REF). Today, it is the extremeness of climate
--- particularly in concert with shifts in mean temperature --- that is
expected to cause the greatest societal damage (REF).

@@

Our results clarify, but do not disagree with previous related analyses.
Previous work with an older version of the GPDD assessed heavy-tailed behaviour
but focused on identifying if the lognormal distribution was the most
frequently parsimonious model of population abundance \citep{halley2002}. For
heavy-tailed distributions, \citet{halley2002} fit the extremely heavy-tailed
Cauchy distribution and the four-parameter Levy stable distribution and found
little information criteria support for these distributions in longer time
series. However, black swan events are by definition rare, and the majority of
time series in the GPDD are short for these purposes. Therefore, we would not
expect to observe black swans in a large proportion of populations.
Furthermore, by probabilistically quantifying process noise tail behaviour on
a continuous spectrum, and allowing for non-stationary time series and density
dependence, our analysis allows for a more nuanced description of the evidence
of ecological black swans. In an earlier study, \citet{keitt1998} described
heavy (power law) tails in breeding bird population abundance. However, this
finding was challenged by \citet{allen2001}, who showed that mixing data across
species could create falsely generate heavy tails. This rebuttal supports our
use of the t distribution to represent heavy tails, since a t distribution can
be represented by a mixture of normal distributions \citep[with
inverse-gamma-distributed variances,][]{gelman2014}.

@@

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.44\textwidth]{nu-coefs-2.pdf}
\caption{
Estimates of how heavy-tailed population dynamics are for \nuCoefPopN\
populations of birds, mammals, insects, and fishes. Small values of $\nu$
($\lesssim 10$) suggest heavy-tailed dynamics; larger values of $\nu$ suggest
approximately normal-tailed dynamics. Vertical points and line segments
represent posterior medians and 50\% / 90\% credible intervals for individual
populations. Inset plots show probability that $\nu < 10$ (probability of heavy
tails) for populations arranged by taxonomic order. Taxonomic orders are sorted
by decreasing mean p($\nu < 10$). Taxonomic orders with three or fewer
populations in panel a are omitted for space. Red and orange colours highlight
populations with a high probability of heavy-tailed dynamics.
}
\label{fig:nu-coefs}
\end{center}
\end{figure}

>>>>

Black swans are unexpected extreme events with dramatic consequences. One of
the most striking black swans in ecology was the asteroid marking the mass
extinction of the K-T boundary (REF). Today, it is the extremeness of climate
--- particularly in concert with shifts in mean temperature --- that is
expected to cause the greatest societal damage (REF).

@@

Our results clarify, but do not disagree with previous related analyses.
Previous work with an older version of the GPDD assessed the distribution of
abundance timeseries but focused on identifying if the lognormal distribution
was the most frequently parsimonious model \citep{halley2002}. For heavy-tailed
distributions, \citet{halley2002} fit the extremely heavy-tailed Cauchy
distribution and the four-parameter Levy stable distribution and found little
information criteria support for these distributions in longer time series.
However, black swan events are by definition rare, and the majority of time
series in the GPDD are short for these purposes. Therefore, we would not expect
to observe black swans in a large proportion of populations. Furthermore, by
probabilistically quantifying process noise tail behaviour on a continuous
spectrum, and allowing for non-stationary time series and density dependence,
our analysis allows for a more nuanced description of the evidence of
ecological black swans. In an earlier study, \citet{keitt1998} described heavy
(power law) tails in breeding bird population abundance. However, this finding
was challenged by \citet{allen2001}, who showed that mixing data across species
could create falsely generate heavy tails. This rebuttal supports our use of
the t distribution to represent heavy tails, since a t distribution can be
represented by a mixture of normal distributions \citep[with
inverse-gamma-distributed variances,][]{gelman2014}.

@@

unlikely to forecast black swans, at least with any useful advanced warning
but we might be able to determine probabilities
and develop policy that is robust to, encourages rapid detection, and that is flexible enought to respond quickly to black swans

example of earch quakes - spatial planning, early warning systems, education systems, a general idea about the probability in given places, build buildings that are robust

Suggests that it's important to have policy that is robust to the uncertainty
about black swans and that can respond and adapt rapidly following black swans

foot and mouth disease and earthquakes?

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.44\textwidth]{nu-coefs-2.pdf}
\caption{
Estimates of how heavy-tailed population dynamics are for \nuCoefPopN\
populations of birds, mammals, insects, and fishes. Small values of $\nu$
($\lesssim 10$) suggest heavy-tailed dynamics; larger values of $\nu$ suggest
approximately normal-tailed dynamics. Vertical points and line segments
represent posterior medians and 50\% / 90\% credible intervals for individual
populations. Inset plots show probability that $\nu < 10$ (probability of heavy
tails) for populations arranged by taxonomic order. Taxonomic orders are sorted
by decreasing mean Pr($\nu < 10$). Taxonomic orders with three or fewer
populations in panel a are omitted for space. Red and orange colours highlight
populations with a high probability of heavy-tailed dynamics.
}
\label{fig:nu-coefs}
\end{center}
\end{figure}
